A template for the \emph{arxiv} style

### 2021-08-10 main/main.Rmd
para on review sf + stars

There has been a large class of implementations dedicated to processing the spatial data. This includes `sf` [@pebesma2018simple] and its precedent `sp` [@pebesma2005s] for ... and `raster` [@raster] and `terra` [@terra] for raster data. While these implementations specialised in geographic manipulations with different type of simple features, it doesn't incorporate a temporal dimension in the data structure. Project like `stars` [@stars] and `spacetime` [@spacetimebook] by R-Spatial allows for both space and time dimension for raster and vector data, but the underlying data structure is a multi-dimensional array, which could be difficult to operate for R users who are more familiar with the operation in 2D dataframe/ tibble.

<!-- - `sp` + `spacetime` (time series data from package `xts`) -->

<!-- - `sf`  -->
<!--   - simple features geometry  -->
<!--   - only spatio -->
<!--   - good for geographical but these geometry types i.e. multipolygon, can be hard to manipulate. Would be nice to have a vector data structure -->

<!-- - `stars` -->
<!--   - a newer structure allow both spatio + temporal  -->
<!--   - array -->

>>>>

Below we review some structure for spatial, temporal, and spatio-temporal data.

Many spatial and spatio-temporal data structures have been developed by the R-spatial team for both raster and vector spatial data. For vector spatial data, which is the focus of this paper, `sf` [@pebesma2018simple] represents spatial vector information with simple features: points, lines, polygons and their multiples. Various `st_` function are designed to manipulate these features based on their geometric relationships. For spatio-temporal data,  `stars` [@stars] can represent both raster and vector data using multi-dimensional array. However, the underlying array structure can be difficult to operate for data analysts who are more familiar with a flat 2D data frame structure used by the tidyverse ecosystem.

### 2021-08-10 main/main.tex
para on review sf + stars

There has been a large class of implementations dedicated to processing
the spatial data. This includes \texttt{sf} (E. J. Pebesma 2018) and its
precedent \texttt{sp} (E. Pebesma and Bivand 2005) for \ldots{} and
\texttt{raster} (Hijmans 2020) and \texttt{terra} (Hijmans 2021) for
raster data. While these implementations specialised in geographic
manipulations with different type of simple features, it doesn't
incorporate a temporal dimension in the data structure. Project like
\texttt{stars} (E. Pebesma 2021) and \texttt{spacetime} (Bivand,
Pebesma, and Gomez-Rubio 2013) by R-Spatial allows for both space and
time dimension for raster and vector data, but the underlying data
structure is a multi-dimensional array, which could be difficult to
operate for R users who are more familiar with the operation in 2D
dataframe/ tibble.

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-spacetimebook}{}%
Bivand, Roger S., Edzer Pebesma, and Virgilio Gomez-Rubio. 2013.
\emph{Applied Spatial Data Analysis with {R}, Second Edition}. Springer,
NY. \url{https://asdar-book.org/}.

\leavevmode\hypertarget{ref-raster}{}%
Hijmans, Robert J. 2020. \emph{Raster: Geographic Data Analysis and
Modeling}. \url{https://CRAN.R-project.org/package=raster}.

\leavevmode\hypertarget{ref-terra}{}%
---------. 2021. \emph{Terra: Spatial Data Analysis}.
\url{https://CRAN.R-project.org/package=terra}.

\leavevmode\hypertarget{ref-stars}{}%
Pebesma, Edzer. 2021. \emph{Stars: Spatiotemporal Arrays, Raster and
Vector Data Cubes}. \url{https://CRAN.R-project.org/package=stars}.

@@

\leavevmode\hypertarget{ref-pebesma2005s}{}%
Pebesma, Edzer, and Roger S Bivand. 2005. {``S Classes and Methods for
Spatial Data: The Sp Package.''} \emph{R News} 5 (2): 9--13.

>>>>

Below we review some structure for spatial, temporal, and
spatio-temporal data.

Many spatial and spatio-temporal data structures have been developed by
the R-spatial team for both raster and vector spatial data. For vector
spatial data, which is the focus of this paper, \texttt{sf} (E. J.
Pebesma 2018) represents spatial vector information with simple
features: points, lines, polygons and their multiples. Various
\texttt{st\_} function are designed to manipulate these features based
on their geometric relationships. For spatio-temporal data,
\texttt{stars} (E. Pebesma 2021) can represent both raster and vector
data using multi-dimensional array. However, the underlying array
structure can be difficult to operate for data analysts who are more
familiar with a flat 2D data frame structure used by the tidyverse
ecosystem.

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-stars}{}%
Pebesma, Edzer. 2021. \emph{Stars: Spatiotemporal Arrays, Raster and
Vector Data Cubes}. \url{https://CRAN.R-project.org/package=stars}.

@@
